Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-26 paper.qmd
section 3.7 and 4 retitled

### Linked plots

Diagnosing the model while locating points in the \gD{} layout and displaying the generated model overlaid on data in the \pD{} is important.

The \gD{} layout and the langevitour view with the model are linked together via rectangular brushes; when a brush is active, points will be highlighted in the adjacent view. Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view. The interface is constructed as a **browsable HTML widget** specifically designed for interactive data analysis.

To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern. It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.

## Assessing the best fit, and hence most accurate \gD{} layout {#sec-bestfit}

>>>>

### Interactive graphics

Matching points in the \gD{} layout with their positions in \pD{} is useful for assessing the fit. This can be used to examine the fitted model in some subspaces in \pD{}, in particular in association with residual plots.

The \gD{} layout and the langevitour view with the fitted model overlaid can be linked using a **browsable HTML widget**. A rectangular "brush" is used to select points in one plot, which will highlight the corresponding points in the other plot(s). Because the langevitour is dynamic, brush events that become active will pause the animation, so that a user can interrogate the current view. This capability will be illustrated on the examples, to show how it can help to understand how the NLDR has organised the observations, and learn where it does not do well.

## Choosing the best \gD{} layout {#sec-bestfit}
